A Critical Appraisal of Zero-Multiple Method in Classical Water Simulation

### 2014-11-03 writedown/zm/paper.water/main.tex
spell checking

Zero-Multiple method, an overview, achievements, sucessful applications.

The propuse of this work: Extensive and strict tests investigating the reliablility of Zero-Multiple method in water simulation, which is the most important solvent.

@@

A short mention of the group cut-off v.s.~the atom cut-off method. (will discuss the implmentation in gromacs later)

@@

In order to investigate the performance of ZM method, we simulation
the water system by using both SPME and ZM method, and calculate a set
of \emph{benchmark quantities}.  This set covers a broad range of
water propertis that are of general interest, and includes the
structure property (radial distribution function), the thermodynamic
properties (pressure, excess chemical potential, constant
volume/pressure heat capacity, isothermal compressibility, thermal
expansion coefficient), dielectric properties (dielectric constant and
Kirkwood-G factor) and dynamical properties (diffusion coefficient and
viscosity).  By comparing ZM to SPME, a good consistency in computing
the benchmark quantities would indicate a good accuracy of the ZM
method.  Moreover, being aware of the weakness of the ZM method also
prevents the application the method in wrong applications.  The ZM
method is also compared to the traditional short-range method: RF.
The definitions for computing these quantities are provided in details
in Appendix.~\ref{appendix:benchmark}.

@@

In this work, we investigate the water system modeled by
TIP3P~\cite{jorgensen1983comparison} point-charge water model.
All simulations were performed by Gromacs~4.6~\cite{hess2008gromacs, pronk2013gromacs} complied by GNU C compiler with single float precision.
The size of the systems and corresponding number of molecules used for MD simulations are summarized in Tab.~\ref{tab:tmp1}.
The benchmark quantities were calculated in either  canonical (NVT) or isothermal-isobaric (NPT) ensemble, with equilibrium temperature $T=300$~K.
In order to simulate at the desired temperature,
the systems $\systemsb$ and $\systemsbp$ were coupled to either Nose-Hoover~\cite{nose1984molecular,hoover1985canonical} thermostat with time-scale $\tau_T = 1.0$~ps,
while systems $\systemmb$, $\systemmbp$, $\systemla$, $\systemlb$ and $\systemlbp$ were coupled to Langevin thermostat with time-scale $\tau_T = 0.1$~ps.
In the NPT simulations, the systems were further coupled
to the Parrinello-Rahman barostat~\cite{parrinello1980crystal,parrinello1981polymorphic} (in Gromacs implementation)
with time-scale $\tau_P = 0.5$~ps to equilibriate the system at 1~Bar.
All systems were integrated with leap-frog scheme at the time-step of 0.001~ps.
For the ZM and SPME methods,
the neighbor lists were updated every 5 time-steps.
The ``\texttt{group}'' cut-off scheme in Gromacs was used. In order to mimic the atomistic
cut-off manner, the radius for neighbor list building was set 0.3~nm
larger than the cut-off radius (the maximum of Coulomb cut-off and the van der Waals cut-off).
To make sure the van der Waals interactions were cut-off in a smooth way,
they are modified in a shell of $0.05$~nm width near the cut-off sphere by the Gromacs ``\texttt{shift}'' method.
The short-range part of SPME method is also smoothed in a shell of $0.05$~nm width  by the Gromacs ``\texttt{pme-switch}'' method.
The reciprocal space
grid spacing is 0.06~nm. The B-Spline interpolation order is 6.
These parameters are chosen to be more accurate than the Gromacs default choices (grid spacing 0.12~nm and interpolation order 4).
The splitting parameter is optimized by the error estimate for the SPME method~\cite{wang2010optimizing}.
The optimized real space and reciprocal space force computing errors are both $1.4\times10^{-4}$~kJ/(mol nm).
In the simulations of the RF method, the interactoins were cut-off in the group manner,
so the radius for neighbor list building is set identical to the cut-off radius.
No cut-off smoothing was applied in the cases of RF simulation.

@@

\begin{figure}
  \centering
  \includegraphics[]{fig/nvt.pressure.1/pressure-methods.eps}
  \caption{The pressure convergence with respect to the cut-off radius.
    The ZM methods with order from 1 to 4 are compared with the RF. 
    The splitting parameter of ZM method is set to be $\alpha = 0.00\,\textrm{nm}^{-1}$.
    The black solid line indicates the maganitude  of the correct pressure: 1~Bar.
    The error bars present the statistical uncertainty at 95\% confidence level.
  }
  \label{fig:pres-comp}
\end{figure}

The equilibrium density of the system was determined by an NPT SPME
simulation.  In the NVT simulation the SPME electrostatic, computed
pressure ($-0.4\pm 1.5$~Bar) is reproducing the target pressure of
1.0~Bar.  The accuracy of ZM method, can be investigate by checking if
the NVT pressure is consistent with the target pressure.  In
Fig.~\ref{fig:pres-comp}, the pressure calculated by ZM method of
order $l=1,2,3$ and 4 are plotted against the cut-off radius.
All the  simulations are done with system $\systemla$.
It is concluded from the Figure that all ZM methods converges to the
target pressure at large enough cut-off radius.
In general, the higher order of ZM presents an larger error in the pressure calculation than
the lower order ZM method.
Higher order are also more sensitve in the  cut-off radius dependency.
ZM $l=1$ tends to under-estimate the pressure, while ZM $l=2,3$ and 4 tends to over-estimate.
The ZM method is also compared with
the RF method using the experimental dielectric constant $\varepsilon_{\textrm{rf}} = 80$, 
from which it is concluded that the accuracy of ZM in pressure calculation is much higher
than the RF method.
It should be noted that even with the largest system of 13824 water molecules,
the fluctuation (measured by standard deviation) of the pressure is roughly 100~Bar, which means that it is
very difficult achieve very accurate estimate for this quantity. In practice a pressure error of 20~Bar is
still acceptable.

\begin{figure}
  \centering
  \includegraphics[width=0.24\textwidth]{fig/nvt.pressure.1/pressure-l1.eps}
  \includegraphics[width=0.24\textwidth]{fig/nvt.pressure.1/pressure-l2.eps}\\
  \includegraphics[width=0.24\textwidth]{fig/nvt.pressure.1/pressure-l3.eps}
  \includegraphics[width=0.24\textwidth]{fig/nvt.pressure.1/pressure-l4.eps}
  \caption{The pressure convergence with respect to the cut-off radius
    for ZM methods with  splitting parameter from $0.0$ to $2.0\,\textrm{nm}^{-1}$.
    From left to right, up to down presensts the ZM method of order 1 to 4.
    The black solid line indicates the maganitude  of the correct pressure: 1~Bar.
    The error bars present the statistical uncertainty at 95\% confidence level.
  }
  \label{fig:pres-l1}
\end{figure}

The effect of using different splitting parameter on the convergence
of the pressure is investigate in Fig.~\ref{fig:pres-l1}.
In general, the higher order indicates less significant dependency on the splitting parameter.
In the case of $l=1$, the splitting parameter 0.0 is consistent with  $0.5\,\textrm{nm}^{-1}$, and
by using higher splitting parameter, the pressure shifts to the higher pressure side.
For $l=2,3$ and 4, the convergence is not sensitive to $\alpha$ upto  $1.0\,\textrm{nm}^{-1}$.
More importantly, for all orders,
the pressure does not converges to the desired value 1~Bar by using $\alpha = 1.5$ or $2.0\,\textrm{nm}^{-1}$.

\begin{sidewaystable}
  \centering
  \caption{The excess chemical potential $\mu^\exc$, constant volume molar heat capacity $C_{v,m}$, dielectric constant $\eps$, diffusion constant $D$, viscosity $\eta$, constant pressure molar heat capacity $C_{v,m}$, isothermal compressibility $\kappa_T$ and
    thermal expansion coefficient $\alpha$
    calculated by different methods.    
    The parenthesises in the last column show the statistical uncertainty
    at the confidence level of 95~\%.
    The bold numbers indicating that its deviation from the SPME result is larger than the statistical uncertainty.
  }
  \bigskip
  \centering\small\setlength\tabcolsep{2pt}
  \begin{tabular*}{0.99\textwidth}{@{\extracolsep{\fill}}cccc cccccccc}\hline\hline
    Method      &   $r_c$ &    $l$ & $\alpha$  & $\mu^\exc$  &$C_{v,m}$ & $C_{p,m}$ &   $\kappa_T$  &$\alpha$ &  $\eps$ & $D$ &  $\eta$  \\
                & [nm] & & [$\textrm{nm}^{-1}$] &   [kJ/mol] &[J/(mol K)] & [J/(mol K)] & [$10^{-10}\textrm{m}^2/\textrm{N}$] &  [$10^{-3}\textrm{K}^{-1}$]&  & [$10^{-9}\textrm{m}^2/\textrm{s}$] &  [$10^{-3}\textrm{Pa}\cdot\textrm{s}$]  \\
                &    &    &    &NVT $\systemla$&NVT $\systemsb$&  NPT $\systemsb$          &  NPT $\systemsb$    &  NPT $\systemsb$    & NVT $\systemsb$& NVT $\systemsb$    & NVT $\systemsb$       \\\hline
    SPME        &1.9 & -- &2.1 & $-26.1$ (0.2) & 72.2 (1.0)  &79.2 (0.9)           & 5.94 (0.05)               &1.03 (0.02)          & 98 (3)          &         5.86  (0.07)&         0.315  (0.007)\\
    RF          &1.2 & -- &--  & $-26.0$ (0.3) & 72.3 (0.8)  &\textbf{82.3} (1.0)  & \textbf{6.51} (0.06)      &\textbf{1.15} (0.02) & \textbf{59} (1) & \textbf{6.27} (0.19)& \textbf{0.449} (0.020)\\\hline
    ZM          &1.2 & 1  &0.0 & $-26.2$ (0.3) & 71.8 (0.9)  &78.9 (1.0)           & 5.87 (0.05)               &1.01 (0.02)          & 96 (2)          & \textbf{5.53} (0.11)& \textbf{0.346} (0.006)\\ 
    ZM          &1.2 & 1  &0.5 & $-26.4$ (0.3) & 72.3 (0.8)  &78.9 (0.9)           & 5.92 (0.05)               &\textbf{1.00} (0.02) & 97 (2)          & \textbf{5.52} (0.25)& \textbf{0.339} (0.008)\\ 
    ZM          &1.2 & 1  &1.0 & $-26.0$ (0.2) & 72.4 (0.8)  &78.4 (0.9)           & 5.95 (0.05)               &1.01 (0.02)          & 96 (2)          &        {5.78} (0.25)& \textbf{0.330} (0.011)\\ 
    ZM          &1.2 & 1  &1.5 & $-26.0$ (0.2) & 71.3 (0.8)  &80.2 (0.9)           & 5.97 (0.05)               &1.04 (0.02)          & 99 (3)          & \textbf{5.71} (0.09)&        {0.312} (0.008)\\ 
    ZM          &1.2 & 1  &2.0 & $-26.1$ (0.3) & 71.3 (0.8)  &79.4 (0.9)           & \textbf{6.05} (0.05)      &1.05 (0.02)          & 96 (3)          &        {5.90} (0.04)&        {0.307} (0.007)\\\hline
    ZM          &1.2 & 2  &0.0 & $-26.2$ (0.2) & 71.5 (0.9)  &79.7 (1.0)           & 5.95 (0.05)               &1.03 (0.02)          & 95 (3)          &         5.82  (0.27)&         0.318  (0.012)\\
    ZM          &1.2 & 3  &0.0 & $-26.2$ (0.2) & 71.6 (0.9)  &79.5 (1.0)           & 5.99 (0.05)               &1.02 (0.02)          & 96 (3)          &         5.79  (0.09)&         0.321  (0.008)\\
    ZM          &1.2 & 4  &0.0 & $-26.0$ (0.3) & 71.1 (0.8)  &79.5 (1.0)           & 5.96 (0.05)               &1.03 (0.02)          &100 (2)          &         5.89  (0.09)&         0.318  (0.013)\\
    \hline\hline
  \end{tabular*}
  \label{tab:thermo}
\end{sidewaystable}

@@

As a measurement of the dipole correlation as a function of molecular distance, the Kirkwood-G factor
is an important qnantity of the water system.
It is well know that the G-factor is system size
dependent~\cite{vanderSpoel2006origin}, therefore we show the G-factor for
both  system $\systemlb$ in FIg.~\ref{fig:gkr}~(a) and a smaller system $\systemmb$
Fig.~\ref{fig:gkr} (b).  In the figures, the statistical uncertainty of
the SPME method is denoted by the red error bars. The statistical uncertainties of the
other methods are essentially of the same size as the SPME, so they are not presented in the figure for
clarity.  In the figure, ZM adaopts a cut-off radius of 1.2~nm and a splitting parameter $\alpha = 0.00\,\textrm{nm}^{-1}$.
RF uses a cut-off 1.2~nm and a dielectric constant of 80.

In both the system, the RF method is qualitatively
wrong. The accuracy of the ZM method is much higher than RF.
In the system of larger size ($\systemlb$), the ZM mehtod of $l=1$ is in consistency with the SPME mehtod. However
order $l=2,3$ and 4  present less but still acceptable accuracy.
In the system of smaller size ($\systemmb$), ZM method of all orders are consistent with the SPME result.
This means ZM performs better in calculated Kirkwood-G factor for smaller systems.
In both systems, ZM $l=1$ presents  an artificial
oscillation around the cut-off radius (1.2~nm), while this oscillation is
not obvious for $l\geq 2$.

\begin{figure}
  \centering
  \includegraphics[]{fig/result.nvt/fig-gkr-conv.eps}
  \includegraphics[]{fig/result.nvt/fig-gkr-l3-damp.eps} 
  \caption{
    The cut-off (a) and spliting parameter (b) dependency in calculating the Kirkwood-G factor.
    The order of ZM method is $l=3$. In plot (a) the spliting parameter is $\alpha = 0.00\,\textrm{nm}^{-1}$.
    in plot (b) the cut-off radius is 1.2~nm.
    The statistical uncertainty is presented at 95~\% confidence level with the  error bars.
  }
  \label{fig:gkr-conv}
\end{figure}

The cut-off and spliting parameter dependency in calculating the Kirkwood-G factor
is presented in Fig.~\ref{fig:gkr-conv}  for ZM method of order $l=3$. With out surprising,
the Kirkwood-G factor converges to the SPME result when using increasing cut-off radius. 
However, unlike the case of calculating RDF, the accuracy Kirkwood-G factor is not improved
with increasing spliting parameter. This fact should be noticed in choosing parameters for
high accuracy calcualtion of Kirkwood-G factor.

@@

The dynamical quantities, diffusion constant and viscosity, are listed in Tab.~\ref{tab:thermo}.
The ZM method of order $l\geq 2$ can almost perfactly reproduce the SPME dynamical quantities.
The ZM order $l=1$ is of less accurate than higher orders, but is more accurate than the RF.
Using larger spliting parameters improves the accuracy, and the value $\alpha= 2.0\ \textrm{nm}^{-1}$ is
sugguested for calculating the dynamical quantities.

@@

In this section we proivde the formula used for computing the benchmark quantities. We assume that there are
$N$ molecules in the water system. The position and configuration of each molecule are fully described
by $\{\vect r_{3i},\vect r_{3i+1}, \vect r_{3i+2}\},\, i\in 1,\cdots,N$, and the corresponding velocities
are denoted by $\{\vect v_{3i},\vect v_{3i+1}, \vect v_{3i+2}\},\, i\in 1,\cdots,N$. The mass of each
atom is denoted by $m_i$.

\subsection{Radial distribution function}
The radial distribution function (RDF) $g(r)$ is a scalor function of distance $r$, which indicates the
probability of finding two atoms of distance $r$ apart.  It is a very
impotant equilibrium structure property, by which the X-ray scattering
intensity and a mount of thermodynamic properties can be computed.
In this work, we investigate the center-of-mass RDF in the water
system. The RDFs presented in the main text were computed in system $\systemlb$.

@@

\subsection{Pressure}
The pressure of the system calculated by the virial formulus:
\begin{align}
  P = \frac1{3V}
  \Bigg\langle
  \sum_{i=1}^{3N} \Big( m_i\vect v_i^2 + \vect r_i\cdot \vect F_i \Big)
  \Bigg\rangle,
\end{align}
where $V$ is the volume of the system. The presented pressures were computed in system $\systemla$.

@@

The chemical potential is defined by
\begin{align}
  \mu = \Big(\frac{\partial A}{\partial N}\Big)_{V,T},
\end{align}
where $A$ is the Helmholtz free energy.
In practice the excess chemical potential $\mu^\exc$, which is the
chemical potential abstracted by the kinetic contribution, is of
special interest. We calculate the excess chemical potential by the
thermodynamic integration (TI). In this approach, the interaction
between an inserted testing molecule and the system is
denotedd by $U_t(\lambda)$, where $\lambda$
is a coupling parameter in range $[0,1]$.
When $\lambda = 1$ the testing molecule is fully coupled to the system
with the TIP3P interaction. When $\lambda=0$, the testing molecule
is decoupled with the system.
Therefore the
Hamiltonian of the system is function of the coupling parameter. The
excess chemical is calculated by the free energy difference between
the fully coupled and decoupled states:
\begin{align}
  \mu^\exc = \int_0^1 \Big\langle \frac{\partial U_t(\lambda)}{\partial \lambda} \Big\rangle_\lambda d\lambda
\end{align}
This integral is usually calculated by numerical integration, which
equally paritions the range of $\lambda$, simulates the system at each
discretized $\lambda$, and computes chemical potential by trapezoidal formulus.
Here for the sake of accuracy, 
we adopt a two step coupling approach: firstly couple the van der Waals
interaction of the testing particle with the rest of the system (stage vdw),
and then couple the electrostatic interaction (stage ele).
During stage vdw, 21 $\lambda$ values are equally
distributed in $[0,1]$.  During stage ele, 6 $\lambda$ values are
equally distributed in $[0,0.05]$, while the other 20 $\lambda$ values
are equally distributed in $(0.05, 1]$. Therefore, in total 47
simulations are performed for all $\lambda$ values.
The free energy differences and the error estimates are calculated
by the Bennet's acceptance ratio method (BAR)~\cite{bennett1976efficient}. 
The simulations were performed with system $\systemla$.

>>>>

Zero-Multiple method, an overview, achievements, successful applications.

The purpose of this work: Extensive and strict tests investigating the reliability of Zero-Multiple method in water simulation, which is the most important solvent.

@@

A short mention of the group cut-off v.s.~the atom cut-off method. (will discuss the implementation in gromacs later)

@@

In order to investigate the performance of ZM method, we simulation
the water system by using both SPME and ZM method, and calculate a set
of \emph{benchmark quantities}.  This set covers a broad range of
water properties that are of general interest, and includes the
structure property (radial distribution function), the thermodynamic
properties (pressure, excess chemical potential, constant
volume/pressure heat capacity, isothermal compressibility, thermal
expansion coefficient), dielectric properties (dielectric constant and
Kirkwood-G factor) and dynamical properties (diffusion coefficient and
viscosity).  By comparing ZM to SPME, a good consistency in computing
the benchmark quantities would indicate a good accuracy of the ZM
method.  Moreover, being aware of the weakness of the ZM method also
prevents the application the method in wrong applications.  The ZM
method is also compared to the traditional short-range method: RF.
The definitions for computing these quantities are provided in details
in Appendix.~\ref{appendix:benchmark}.

@@

In this work, we investigate the water system modeled by
TIP3P~\cite{jorgensen1983comparison} point-charge water model.
All simulations were performed by Gromacs~4.6~\cite{hess2008gromacs, pronk2013gromacs} complied by GNU C compiler with single float precision.
The size of the systems and corresponding number of molecules used for MD simulations are summarized in Tab.~\ref{tab:tmp1}.
The benchmark quantities were calculated in either  canonical (NVT) or isothermal-isobaric (NPT) ensemble, with equilibrium temperature $T=300$~K.
In order to simulate at the desired temperature,
the systems $\systemsb$ and $\systemsbp$ were coupled to either Nose-Hoover~\cite{nose1984molecular,hoover1985canonical} thermostat with time-scale $\tau_T = 1.0$~ps,
while systems $\systemmb$, $\systemmbp$, $\systemla$, $\systemlb$ and $\systemlbp$ were coupled to Langevin thermostat with time-scale $\tau_T = 0.1$~ps.
In the NPT simulations, the systems were further coupled
to the Parrinello-Rahman barostat~\cite{parrinello1980crystal,parrinello1981polymorphic} (in Gromacs implementation)
with time-scale $\tau_P = 0.5$~ps to equilibriate the system at 1~Bar.
All systems were integrated with leap-frog scheme at the time-step of 0.001~ps.
For the ZM and SPME methods,
the neighbor lists were updated every 5 time-steps.
The ``\texttt{group}'' cut-off scheme in Gromacs was used. In order to mimic the atomistic
cut-off manner, the radius for neighbor list building was set 0.3~nm
larger than the cut-off radius (the maximum of Coulomb cut-off and the van der Waals cut-off).
To make sure the van der Waals interactions were cut-off in a smooth way,
they are modified in a shell of $0.05$~nm width near the cut-off sphere by the Gromacs ``\texttt{shift}'' method.
The ZM method was implemented in Gromacs with the tabulated electrostatic interaction (Home-made codes for generating the interaction tables).
The short-range part of SPME method is also smoothed in a shell of $0.05$~nm width  by the Gromacs ``\texttt{pme-switch}'' method.
The reciprocal space
grid spacing is set to 0.06~nm, with B-Spline interpolation order of 6.
These parameters are chosen to be more accurate than the Gromacs default choices (grid spacing 0.12~nm and interpolation order 4).
The splitting parameter is optimized by the error estimate for the SPME method, which is implemented in Gromacs by tool ``\texttt{g\_pme\_error}''~\cite{wang2010optimizing}.
The optimized real space and reciprocal space force computing errors were equally distributed as $1.4\times10^{-4}$~kJ/(mol nm).
In the simulations of the RF method, the interactions were cut-off in the group manner,
so the radius for neighbor list building is set identical to the cut-off radius.
No cut-off smoothing was applied in the cases of RF simulation.

@@

\begin{figure}
  \centering
  \includegraphics[]{fig/nvt.pressure.1/pressure-methods.eps}
  \caption{The pressure convergence with respect to the cut-off radius.
    The ZM methods with order from 1 to 4 are compared with the RF. 
    The splitting parameter of ZM method is set to be $\alpha = 0.00\,\textrm{nm}^{-1}$.
    The black solid line indicates the magnitude  of the correct pressure: 1~Bar.
    The error bars present the statistical uncertainty at 95\% confidence level.
  }
  \label{fig:pres-comp}
\end{figure}

The equilibrium density of the system was determined by an NPT SPME
simulation.  In the NVT simulation the SPME electrostatic, computed
pressure ($-0.4\pm 1.5$~Bar) is reproducing the target pressure of
1.0~Bar.  The accuracy of ZM method, can be investigate by checking if
the NVT pressure is consistent with the target pressure.  In
Fig.~\ref{fig:pres-comp}, the pressure calculated by ZM method of
order $l=1,2,3$ and 4 are plotted against the cut-off radius.
All the  simulations are done with system $\systemla$.
It is concluded from the Figure that all ZM methods converges to the
target pressure at large enough cut-off radius.
In general, the higher order of ZM presents an larger error in the pressure calculation than
the lower order ZM method.
Higher order are also more sensitive in the  cut-off radius dependency.
ZM $l=1$ tends to under-estimate the pressure, while ZM $l=2,3$ and 4 tends to over-estimate.
The ZM method is also compared with
the RF method using the experimental dielectric constant $\varepsilon_{\textrm{rf}} = 80$, 
from which it is concluded that the accuracy of ZM in pressure calculation is much higher
than the RF method.
It should be noted that even with the largest system of 13824 water molecules,
the fluctuation (measured by standard deviation) of the pressure is roughly 100~Bar, which means that it is
very difficult achieve very accurate estimate for this quantity. In practice a pressure error of 20~Bar is
still acceptable.

\begin{figure}
  \centering
  \includegraphics[width=0.24\textwidth]{fig/nvt.pressure.1/pressure-l1.eps}
  \includegraphics[width=0.24\textwidth]{fig/nvt.pressure.1/pressure-l2.eps}\\
  \includegraphics[width=0.24\textwidth]{fig/nvt.pressure.1/pressure-l3.eps}
  \includegraphics[width=0.24\textwidth]{fig/nvt.pressure.1/pressure-l4.eps}
  \caption{The pressure convergence with respect to the cut-off radius
    for ZM methods with  splitting parameter from $0.0$ to $2.0\,\textrm{nm}^{-1}$.
    From left to right, up to down presents the ZM method of order 1 to 4.
    The black solid line indicates the magnitude  of the correct pressure: 1~Bar.
    The error bars present the statistical uncertainty at 95\% confidence level.
  }
  \label{fig:pres-l1}
\end{figure}

The effect of using different splitting parameter on the convergence
of the pressure is investigate in Fig.~\ref{fig:pres-l1}.
In general, the higher order indicates less significant dependency on the splitting parameter.
In the case of $l=1$, the splitting parameter 0.0 is consistent with  $0.5\,\textrm{nm}^{-1}$, and
by using higher splitting parameter, the pressure shifts to the higher pressure side.
For $l=2,3$ and 4, the convergence is not sensitive to $\alpha$ up to  $1.0\,\textrm{nm}^{-1}$.
More importantly, for all orders,
the pressure does not converges to the desired value 1~Bar by using $\alpha = 1.5$ or $2.0\,\textrm{nm}^{-1}$.

\begin{sidewaystable}
  \centering
  \caption{The excess chemical potential $\mu^\exc$, constant volume molar heat capacity $C_{v,m}$, dielectric constant $\eps$, diffusion constant $D$, viscosity $\eta$, constant pressure molar heat capacity $C_{v,m}$, isothermal compressibility $\kappa_T$ and
    thermal expansion coefficient $\alpha$
    calculated by different methods.    
    The parentheses in the last column show the statistical uncertainty
    at the confidence level of 95~\%.
    The bold numbers indicating that its deviation from the SPME result is larger than the statistical uncertainty.
  }
  \bigskip
  \centering\small\setlength\tabcolsep{2pt}
  \begin{tabular*}{0.99\textwidth}{@{\extracolsep{\fill}}cccc cccccccc}\hline\hline
    Method      &   $r_c$ &    $l$ & $\alpha$  & $\mu^\exc$  &$C_{v,m}$ & $C_{p,m}$ &   $\kappa_T$  &$\alpha$ &  $\eps$ & $D$ &  $\eta$  \\
                & [nm] & & [$\textrm{nm}^{-1}$] &   [kJ/mol] &[J/(mol K)] & [J/(mol K)] & [$10^{-10}\textrm{m}^2/\textrm{N}$] &  [$10^{-3}\textrm{K}^{-1}$]&  & [$10^{-9}\textrm{m}^2/\textrm{s}$] &  [$10^{-3}\textrm{Pa}\cdot\textrm{s}$]  \\
                &    &    &    &NVT $\systemla$&NVT $\systemsb$&  NPT $\systemsb$          &  NPT $\systemsb$    &  NPT $\systemsb$    & NVT $\systemsb$& NVT $\systemsb$    & NVT $\systemsb$       \\\hline
    SPME        &1.9 & -- &2.1 & $-26.1$ (0.2) & 72.2 (1.0)  &79.2 (0.9)           & 5.94 (0.05)               &1.03 (0.02)          & 98 (3)          &         5.86  (0.07)&         0.315  (0.007)\\
    RF          &1.2 & -- &--  & $-26.0$ (0.3) & 72.3 (0.8)  &\textbf{82.3} (1.0)  & \textbf{6.51} (0.06)      &\textbf{1.15} (0.02) & \textbf{59} (1) & \textbf{6.27} (0.19)& \textbf{0.449} (0.020)\\\hline
    ZM          &1.2 & 1  &0.0 & $-26.2$ (0.3) & 71.8 (0.9)  &78.9 (1.0)           & 5.87 (0.05)               &1.01 (0.02)          & 96 (2)          & \textbf{5.53} (0.11)& \textbf{0.346} (0.006)\\ 
    ZM          &1.2 & 1  &0.5 & $-26.4$ (0.3) & 72.3 (0.8)  &78.9 (0.9)           & 5.92 (0.05)               &\textbf{1.00} (0.02) & 97 (2)          & \textbf{5.52} (0.25)& \textbf{0.339} (0.008)\\ 
    ZM          &1.2 & 1  &1.0 & $-26.0$ (0.2) & 72.4 (0.8)  &78.4 (0.9)           & 5.95 (0.05)               &1.01 (0.02)          & 96 (2)          &        {5.78} (0.25)& \textbf{0.330} (0.011)\\ 
    ZM          &1.2 & 1  &1.5 & $-26.0$ (0.2) & 71.3 (0.8)  &80.2 (0.9)           & 5.97 (0.05)               &1.04 (0.02)          & 99 (3)          & \textbf{5.71} (0.09)&        {0.312} (0.008)\\ 
    ZM          &1.2 & 1  &2.0 & $-26.1$ (0.3) & 71.3 (0.8)  &79.4 (0.9)           & \textbf{6.05} (0.05)      &1.05 (0.02)          & 96 (3)          &        {5.90} (0.04)&        {0.307} (0.007)\\\hline
    ZM          &1.2 & 2  &0.0 & $-26.2$ (0.2) & 71.5 (0.9)  &79.7 (1.0)           & 5.95 (0.05)               &1.03 (0.02)          & 95 (3)          &         5.82  (0.27)&         0.318  (0.012)\\
    ZM          &1.2 & 3  &0.0 & $-26.2$ (0.2) & 71.6 (0.9)  &79.5 (1.0)           & 5.99 (0.05)               &1.02 (0.02)          & 96 (3)          &         5.79  (0.09)&         0.321  (0.008)\\
    ZM          &1.2 & 4  &0.0 & $-26.0$ (0.3) & 71.1 (0.8)  &79.5 (1.0)           & 5.96 (0.05)               &1.03 (0.02)          &100 (2)          &         5.89  (0.09)&         0.318  (0.013)\\
    \hline\hline
  \end{tabular*}
  \label{tab:thermo}
\end{sidewaystable}

@@

As a measurement of the dipole correlation as a function of molecular distance, the Kirkwood-G factor
is an important quantity of the water system.
It is well know that the G-factor is system size
dependent~\cite{vanderSpoel2006origin}, therefore we show the G-factor for
both  system $\systemlb$ in Fig.~\ref{fig:gkr}~(a) and a smaller system $\systemmb$
Fig.~\ref{fig:gkr} (b).  In the figures, the statistical uncertainty of
the SPME method is denoted by the red error bars. The statistical uncertainties of the
other methods are essentially of the same size as the SPME, so they are not presented in the figure for
clarity.  In the figure, ZM adopts a cut-off radius of 1.2~nm and a splitting parameter $\alpha = 0.00\,\textrm{nm}^{-1}$.
RF uses a cut-off 1.2~nm and a dielectric constant of 80.

In both the system, the RF method is qualitatively
wrong. The accuracy of the ZM method is much higher than RF.
In the system of larger size ($\systemlb$), the ZM method of $l=1$ is in consistency with the SPME method. However
order $l=2,3$ and 4  present less but still acceptable accuracy.
In the system of smaller size ($\systemmb$), ZM method of all orders are consistent with the SPME result.
This means ZM performs better in calculated Kirkwood-G factor for smaller systems.
In both systems, ZM $l=1$ presents  an artificial
oscillation around the cut-off radius (1.2~nm), while this oscillation is
not obvious for $l\geq 2$.

\begin{figure}
  \centering
  \includegraphics[]{fig/result.nvt/fig-gkr-conv.eps}
  \includegraphics[]{fig/result.nvt/fig-gkr-l3-damp.eps} 
  \caption{
    The cut-off (a) and splitting parameter (b) dependency in calculating the Kirkwood-G factor.
    The order of ZM method is $l=3$. In plot (a) the splitting parameter is $\alpha = 0.00\,\textrm{nm}^{-1}$.
    in plot (b) the cut-off radius is 1.2~nm.
    The statistical uncertainty is presented at 95~\% confidence level with the  error bars.
  }
  \label{fig:gkr-conv}
\end{figure}

The cut-off and splitting parameter dependency in calculating the Kirkwood-G factor
is presented in Fig.~\ref{fig:gkr-conv}  for ZM method of order $l=3$. With out surprising,
the Kirkwood-G factor converges to the SPME result when using increasing cut-off radius. 
However, unlike the case of calculating RDF, the accuracy Kirkwood-G factor is not improved
with increasing splitting parameter. This fact should be noticed in choosing parameters for
high accuracy calculation of Kirkwood-G factor.

@@

The dynamical quantities, diffusion constant and viscosity, are listed in Tab.~\ref{tab:thermo}.
The ZM method of order $l\geq 2$ can almost perfectly reproduce the SPME dynamical quantities.
The ZM order $l=1$ is of less accurate than higher orders, but is more accurate than the RF.
Using larger splitting parameters improves the accuracy, and the value $\alpha= 2.0\ \textrm{nm}^{-1}$ is
suggested for calculating the dynamical quantities.

@@

In this section we provide the formula used for computing the benchmark quantities. We assume that there are
$N$ molecules in the water system. The position and configuration of each molecule are fully described
by $\{\vect r_{3i},\vect r_{3i+1}, \vect r_{3i+2}\},\, i\in 1,\cdots,N$, and the corresponding velocities
are denoted by $\{\vect v_{3i},\vect v_{3i+1}, \vect v_{3i+2}\},\, i\in 1,\cdots,N$. The mass of each
atom is denoted by $m_i$.

\subsection{Radial distribution function}
The radial distribution function (RDF) $g(r)$ is a scalar function of distance $r$, which indicates the
probability of finding two atoms of distance $r$ apart.  It is a very
important equilibrium structure property, by which the X-ray scattering
intensity and a mount of thermodynamic properties can be computed.
In this work, we investigate the center-of-mass RDF in the water
system. The RDFs presented in the main text were computed in system $\systemlb$.

@@

\subsection{Pressure}
The pressure of the system calculated by the virial formula:
\begin{align}
  P = \frac1{3V}
  \Bigg\langle
  \sum_{i=1}^{3N} \Big( m_i\vect v_i^2 + \vect r_i\cdot \vect F_i \Big)
  \Bigg\rangle,
\end{align}
where $V$ is the volume of the system. The presented pressures were computed in system $\systemla$.

@@

The chemical potential is defined by
\begin{align}
  \mu = \Big(\frac{\partial A}{\partial N}\Big)_{V,T},
\end{align}
where $A$ is the Helmholtz free energy.
In practice the excess chemical potential $\mu^\exc$, which is the
chemical potential abstracted by the kinetic contribution, is of
special interest. We calculate the excess chemical potential by the
thermodynamic integration (TI). In this approach, the interaction
between an inserted testing molecule and the system is
denoted by $U_t(\lambda)$, where $\lambda$
is a coupling parameter in range $[0,1]$.
When $\lambda = 1$ the testing molecule is fully coupled to the system
with the TIP3P interaction. When $\lambda=0$, the testing molecule
is decoupled with the system.
Therefore the
Hamiltonian of the system is function of the coupling parameter. The
excess chemical is calculated by the free energy difference between
the fully coupled and decoupled states:
\begin{align}
  \mu^\exc = \int_0^1 \Big\langle \frac{\partial U_t(\lambda)}{\partial \lambda} \Big\rangle_\lambda d\lambda
\end{align}
This integral is usually calculated by numerical integration, which
equally partitions the range of $\lambda$, simulates the system at each
discretized $\lambda$, and computes chemical potential by trapezoidal formula.
Here for the sake of accuracy, 
we adopt a two step coupling approach: firstly couple the van der Waals
interaction of the testing particle with the rest of the system (stage vdw),
and then couple the electrostatic interaction (stage ele).
During stage vdw, 21 $\lambda$ values are equally
distributed in $[0,1]$.  During stage ele, 6 $\lambda$ values are
equally distributed in $[0,0.05]$, while the other 20 $\lambda$ values
are equally distributed in $(0.05, 1]$. Therefore, in total 47
simulations are performed for all $\lambda$ values.
The free energy differences and the error estimates are calculated
by the Bennet's acceptance ratio method (BAR)~\cite{bennett1976efficient}. 
The simulations were performed with system $\systemla$.
